---
title: "Simulation"
editor_options: 
  chunk_output_type: console
execute:
  echo: false
---

```{r, include = FALSE}
library(tidyverse)
library(fec16)
library(skimr)
library(ggplot2)
library(ggspatial)
library(sf)
library(stringr)
library(spdep)
library(spatstat)
```

```{r}
encampments <- read_csv("Data/homelessness.csv")
encampments <- encampments |>
  mutate(
    Latitude = as.numeric(Latitude),
    Longitude = as.numeric(Longitude)
  ) |> 
  filter(!is.na(Latitude)) |> 
  filter(!is.na(Longitude))

encampments_sf <- st_as_sf(encampments, coords = c("Longitude", "Latitude"), remove = FALSE)
encampments_sf <- encampments_sf |> st_set_crs(4326)

nyc_sf <- st_read("Data/Borough Boundaries/geo_export_b711edb6-f22b-4799-a1ca-fb9c325f7f7e.shp") |> st_transform(4326)

mn_sf <- st_read("Data/2010 Neighborhood Tabulation Areas (NTAs)/geo_export_e68e4089-10c6-4448-8f19-7c31293a1f7a.shp") |> st_transform(4326) |> 
  filter(boro_name == "Manhattan")
```


```{r}
ggplot() +
  geom_sf(data = nyc_sf) +
  geom_sf(data = encampments_sf) 
```

This map shows the spread of service requests called on homeless encampments in NYC between 2000 until March 22, 2024. While the COVID pandemic had a significant impact on the number of encampments in the city, the data is helpful in understanding relatively high areas of homelessness in the city. It is clear that Manhattan has the highest number of reported encampments, so let's zoom in on it.


```{r}
camps_mn <- encampments_sf |> 
  filter(Borough == "MANHATTAN")

ggplot() +
  geom_sf(data = mn_sf) +
  geom_sf(data = camps_mn)
```

We can see that some neighborhoods have had many reports of camps and some none at all. If we aggregate the data we will have a better time visually assessing any spatial autocorrelation.

```{r}
mn_sf <- mn_sf |> 
  mutate(n = lengths(st_intersects(mn_sf, camps_mn)),
         area = st_area(mn_sf))
```


```{r}
camps_by_neighborhood <- camps_mn |> 
  group_by(`Community Districts`) |> 
  summarise(
    n = n()
  ) |> as.data.frame()
```



```{r}
for (i in seq_along(mn_sf$n)) {
  if (is.na(mn_sf$n[i])) {
    mn_sf$n[i] <- 0
  }
}

mn_sf <- mn_sf |> mutate_at("ntacode", str_replace, "MN", "") |> 
  mutate(
    ntacode = as.numeric(ntacode)
  )

map_fill <- function(dataset, variable) { # Function
  dataset |> 
    ggplot() +
    geom_sf(aes(fill = variable)) +
    scale_fill_viridis_c()
}

map_fill(mn_sf, mn_sf$n)
```

We can see that many community districts in Manhattan have had little to no reports of encampments, while others have a concentrated number. We can better visualise the data if we log the number of encampments.

```{r}
map_fill(mn_sf, log(mn_sf$n))
```

Much better, we can better visually assess the distribution of reported encampments in the community districts.

From here, we can calculate the Moran's I which is a statistical test that measures the spatial autocorrelation of an aggregate variable.

```{r}
set.seed(1)
mn_nb <- poly2nb(mn_sf, queen = FALSE) 
mn_nbw <- nb2listw(mn_nb, style = "W", zero.policy = TRUE)
moran.mc(mn_sf$n, mn_nbw, nsim = 999) # Loop
```

$H_O$: No spatial autocorrelation, I is close to 0

$H_A$: Spatial autocorrelation, I $\neq$ 0. 

With a Moran's I of -0.013, we observe a weak negative spatial autocorrelation. However, as the p-value of 0.299 is relatively large, we fail to reject the null hypothesis; the number of reported homeless encampments is not spatially autocorrelated in Manhattan community districts as per this data and our analysis. This could be due to many factors, the lack of robust data on homelessness. Temporal bias could also be an issue, as the data is spread out over four years. Aggregate bias could also have an effect on the results; analysis of point-pattern data using K-cross analyses would have been more appropriate.


